Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-19 papers/clusters.tex
Typos, question for Hogg, and signal-to-noise --> /N$ post-definition

Part of the reason that chemical tagging still remains unrealized
is because the level of abundance specificity required
is very high.
If there are thousands of (relevant) molecular clouds forming stars in
the recent history of the Milky Way, clumps of stars can only be
identified in abundance space if abundance space is high in dimensionality.
(In principle, it needs to be high in dimensionality both in terms
of the number of measured abundances \emph{and} in terms of the number of
neucleosynthetic pathways or the dimensionality of the true abundance
space.)
Therefore accurate---or at
least precise---measurements of many different abundances are needed
for stellar sibling to have sufficiently unique fingerprints.
Stellar spectroscopic surveys now have the resolution,
signal-to-noise, wavelength coverage, and sample sizes to deliver many
different chemical tags for each star.

@@

The second issue for chemical tagging---and the one we address
here---is measurement precision (and accuracy).
The current precision on abundance measurements in the published
survey catalogs is not high enough (for example, \citealt{martell,
  ting}).
However, our recent work \citep{thecannon} suggests that \emph{data}
are precise enough: there is enough signal-to-noise at the relevant
locations in spectrum space to deliver high-precision tags.
The existence of very large data sets, homogeneous in data/spectrum
space, suggest that one can use data-driven approaches to determining
stellar abundances, which can considerably outperform traditional
methods.
These traditional approaches are based on \foreign{ab-initio} physical models
that have shortcomings that become apparent in this age of
high-quality spectra, and the data-analysis methods do not make use of
all of the information in the data sets.
Improving the models and exploiting the entire information
content in the data is critical if we are going to deliver useful
chemical tags.

@@

It is important to realize that analysis techniques based on \foreign{ab-initio}
physical models are inaccurate, yet they too strive to improve precision
in the face of knowingly inaccurate models.  Incomplete atomic data, 
simplifications of photospheric structure, presumptions about convective
motion, and inconsistencies resulting from assumptions of the local thermal
equilibrium, all contribute to produce knowingly inaccurate abundances.
\thecannon\ stands out because it demonstrably improves the precision on
chemical abundances, whereas the accuracy of those labels is limited only
by the training set employed.

@@

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{dr12}, alll those not
flagged as bad in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because this
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here.
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) form the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$ objects the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels. In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the model coefficients fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is embarrassingly parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.

>>>>

Part of the reason that chemical tagging still remains unrealized
is because the level of abundance specificity required
is very high.
If there are thousands of (relevant) molecular clouds forming stars in
the recent history of the Milky Way, clumps of stars can only be
identified in abundance space if abundance space is high in dimensionality.
(In principle, it needs to be high in dimensionality both in terms
of the number of measured abundances \emph{and} in terms of the number of
nucleosynthetic pathways or the dimensionality of the true abundance
space.)
Therefore accurate---or at
least precise---measurements of many different abundances are needed
for stellar sibling to have sufficiently unique fingerprints.
Stellar spectroscopic surveys now have the resolution,
$S/N$, wavelength coverage, and sample sizes to deliver many
different chemical tags for each star.

@@

The second issue for chemical tagging---and the one we address
here---is measurement precision (and accuracy).
The current precision on abundance measurements in the published
survey catalogs is not high enough (for example, \citealt{martell,
  ting}).
However, our recent work \citep{thecannon} suggests that \emph{data}
are precise enough: there is enough $S/N$ at the relevant
locations in spectrum space to deliver high-precision tags.
The existence of very large data sets, homogeneous in data/spectrum
space, suggest that one can use data-driven approaches to determining
stellar abundances, which can considerably outperform traditional
methods.
These traditional approaches are based on \foreign{ab-initio} physical models
that have shortcomings that become apparent in this age of
high-quality spectra, and the data-analysis methods do not make use of
all of the information in the data sets.
Improving the models and exploiting the entire information
content in the data is critical if we are going to deliver useful
chemical tags.

@@

It is important to realize that analysis techniques based on \foreign{ab-initio}
physical models are inaccurate, yet they too strive to improve precision
in the face of knowingly inaccurate models.  Incomplete atomic data, 
simplifications of photospheric structure, presumptions about convective
motion, and inconsistencies resulting from assumptions of the local thermal
equilibrium all contribute to produce knowingly inaccurate abundances.
\thecannon\ stands out because it demonstrably improves the precision on
chemical abundances, whereas the accuracy of those labels is limited only
by the training set employed.

@@

% ARC asks: "not flagged as bad" --> what exactly do we mean here? ASPCAPFLAG = 0?

Our analysis draws on the spectra of \totalnumber\ giant stars ($\logg
< 3.9$) from \apogee\ \acronym{DR12} \citep{dr12}, all those not
flagged as bad in the \apogee\ \aspcap\ \citep{aspcap} pipeline
reductions.
We re-analyze these \emph{spectra} using \thecannon, because this
deliver stellar abundance labels of higher precision, especially for
stars of $S/N \le 150$.
The details about how we select, reduce, and analyze the \apogee\ data
are given in full detail in the companion papers \citep{casey16,
  ness16}, and we only summarize briefly here.
We re-derived 17 labels ($\teff$, $\logg$, and 15 abundances
referenced to H) form the \apogee\ \acronym{DR12} spectra. 
For the training step of \thecannon, 
12,681 red-giant stars with spectral $S/N \ge 200$ were used.
For this small fraction of the sample with the highest $S/N$ objects the
labels provided by \aspcap\ provide consistent, relatively low-scatter, sensible
abundance-space measurements.
The training step fixes a spectral model
that predicts the normalized spectrum as a quadratic function of all
labels. In the second stage---the test step---each unlabelled star is
used to establish a single-star likelihood function for its labels,
holding the model coefficients fixed.
\thecannon\ finds the labels that minimize the single-star likelihood
function for each test-set star.
This optimization is not convex, but it is embarrassingly parallel,
and therefore fast. The test
step for 150,000 spectra takes less than 30~minutes on a small 
research cluster in Cambridge.
